---
title: "Biking @ Penn"
author: "Jake"
date: "June 23, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F)
```

```{r}
setwd("J:/Portfolio/Biking@Penn")
library(tidyverse)
library(stringr)
library(ggmap)

get_intersections <- read_csv("get_intersections.csv")
get_centroids <- read_csv("get_centroids.csv")
get_routes <- read_csv(make_routes, "make_routes.csv")
```

```{r}
ggplot() +
    geom_point(data = get_centroids, aes(X,Y), color = "blue") +
    geom_point(data = get_intersections, aes(X,Y), color = "orange") +
    theme_classic()
```

```{r}
prep_routes <-
    make_routes %>% 
    left_join(get_centroids, by = c("centroid" = "id"))
    group_by()
    
    

```


to get polylines, look at bottom answer: `googleway` package

https://stackoverflow.com/questions/30270011/ggmap-route-finding-doesnt-stay-on-roads





```{r}
a <-# loc_pair "16 19"
    route(from = "39.9409445009992,-75.1976984808205",
      to = "39.9549319483263,-75.1930786028087",
      mode = "driving",
      structure = "route",
      output = "all")

b <-
    route(from = "39.948614,-75.2184281",
      to = "39.9522327,-75.1947939",
      mode = "walking",
      structure = "route")

ggplot() +
    geom_path(data = a, aes(x = lon, y = lat), color = "red", size = 2) +
    geom_path(data = b, aes(x = lon, y = lat), size = 1)

a1 <- 
    make_routes %>% 
    filter(route == "08 29", type == "driving")

ggplot() +
    geom_path(data = prep_routes 
              #%>% slice(1:100)
              , 
              aes(x = lon, 
                  y = lat, 
                  group = route, 
                  color = type), 
              alpha = .2,
              size = 1) +
    #geom_text(data = get_intersections, aes(X,Y, label = id), size = 8) +
    #geom_text(data = get_centroids, aes(X,Y, label = id), color = "blue") +
    theme_classic()

sum(a$miles, na.rm = T)
sum(b$miles, na.rm = T)
sum(a$minutes, na.rm = T)
sum(b$minutes, na.rm = T)

routeQueryCheck()


?route


a$routes[[1]]$overview_polyline

a$routes[[1]]$legs[[1]]$steps[[1]]$polyline
```

#get csv's from shp
```{r make_csv, eval = F}
library(sf)
get_intersections <-
    st_read("penn_street_intersections.shp") %>% 
    #filter(Direction != "W") %>% 
    cbind(st_coordinates(.)) %>% 
    mutate(id = row_number() %>% str_pad(., 2, "l","0"),
           start = paste(Y, X, sep = ",")) %>% 
    st_set_geometry(NULL)

get_centroids <-
    st_read("penn_building_centroids.shp") %>% 
    cbind(st_coordinates(.)) %>% 
    mutate(id = row_number() %>% str_pad(., 2, "l","0"),
           end = paste(Y, X, sep = ",")) %>% 
    st_set_geometry(NULL)

write_csv(get_intersections, "get_intersections.csv")
write_csv(get_centroids, "get_centroids.csv")
```

#data prep
```{r prep_list, eval = F}
get_directions <- 
    expand.grid(
        intersection = get_intersections$id,
        centroid = get_centroids$id,
        type = c("driving", "walking"),
        stringsAsFactors = F,
        KEEP.OUT.ATTRS = F) %>%
    left_join(get_intersections %>% 
                  select(id, start), by = c("intersection" = "id")) %>%
    left_join(get_centroids %>% 
                  select(id, end), by = c("centroid" = "id")) %>%
    mutate(loc_pair = paste(intersection, centroid),
           route = paste(loc_pair, type))
```

```{r loop_function, eval = F}
get_route <- 
    function(i) {
        route(from = get_directions$start[i],
              to = get_directions$end[i], 
              mode = get_directions$type[i], 
              structure = "route") %>%
        mutate(loc_pair = get_directions$loc_pair[i],
               route = get_directions$route[i],
               type = get_directions$type[i],
               intersection = get_directions$intersection[i],
               centroid = get_directions$centroid[i])
    }

make_routes <- get_route(1)

```

#ggmap routes loop
```{r loop, eval = F}
time_diff <-
    function(x){
        difftime(Sys.time(), x, units = "secs") %>% as.numeric() %>% floor()
    }

time_start <- Sys.time()

for(i in 2:nrow(get_directions)){

    make_routes <- rbind(make_routes, get_route(i))
    
    Sys.sleep(0.5)
    
    print(paste(i, "-", time_diff(time_start), 'seconds'))
}


routeQueryCheck() #check to see how much of quota is used (2500)

write_csv(make_routes, "make_routes.csv")

```


```{r}
library(googleway)
library(geosphere)

set_key(key = key)
google_keys()


get_route <- 
    function(i) {
        route_list <-
         google_directions(origin = get_directions$start[i],
                          destination = get_directions$end[i],
                          #key = mykey,
                          mode = get_directions$type[i], 
                          avoid = "highways",
                          simplify = TRUE)
        
    #return dataframe:
        decode_pl(route_list$routes$overview_polyline$points) %>% 
        mutate(loc_pair = get_directions$loc_pair[i],
               route = get_directions$route[i],
               type = get_directions$type[i],
               intersection = get_directions$intersection[i],
               centroid = get_directions$centroid[i],
               ord = row_number())
    }

make_routes <- get_route(1)

time_diff <-
    function(x){
        difftime(Sys.time(), x, units = "secs") %>% 
            as.numeric() %>% 
            floor()
    }

for(i in 2:nrow(get_directions)){

    make_routes <- rbind(make_routes, get_route(i))
    
    Sys.sleep(0.5)
    
    print(paste(i, "-", time_diff(time_start), 'seconds'))
}

time_diff(time_start) - 321303

write_csv(make_routes, "make_routes_googleway.csv")



route_distances <-
    make_routes %>%
    #slice(1:1000) %>% 
    group_by(route) %>% 
    mutate(lat2 = ifelse(ord == max(ord), lat, lead(lat)),
           lon2 = ifelse(ord == max(ord), lon, lead(lon))) %>% 
    ungroup() %>% 
    rowwise() %>%
    mutate(dist = distHaversine(c(lon, lat), c(lon2, lat2), r = 3958.756)) %>% 
    ungroup()

write_csv(route_distances, "make_routes_googleway.csv")

route_agg <-
    route_distances %>% 
    group_by(loc_pair, route, type, intersection) %>% 
    summarise(dist = sum(dist)) %>% 
    group_by(loc_pair) %>% 
    mutate(abs_diff = ifelse(dist == min(dist), 0, max(dist)-min(dist))) %>% 
    ungroup()

write_csv(route_agg, "route_agg.csv")


ggplot(make_routes, aes(lon, lat, label = ord)) +
    geom_path() +
    geom_text() +
    coord_sf()



