---
title: "Flexdashboard Demo"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    source_code: embed
    theme: lumen
    vertical_layout: fill
runtime: shiny
---

```{r input_demo, eval = FALSE}
# this chunk doesn't run but works to troubleshoot reactive data 
input <- list(
  displ = c(2, 5),
  year = 2008
)
```

```{r setup}
library(tidyverse)
library(shiny)        #renderPlot, renderTable, sliderInput
library(shinyWidgets) #radioGroupButtons
raw_data <- mpg 


# the layout of a flexdashboard can be controlled with ====== or ----- to separate sections
# ===== will create new pages
# ----- will create new containers and are accompanied by Column or Row
# the {} in these sections control parameters such as width, class, tabs or hidden status
# the use of ### create sub containers to hold plots, tables, etc
```

MPG Example
=====================================
Column {.sidebar data-width=200} 
-------------------------------------
Filter the data:

```{r filters}
# Line 36 makes this a sidebar. 
# You can make this sidebar global by putting this chunk abovre MPG Example (line 34)
# below are 2 input options: a group of buttons, and a slider

# many widgets take a choices argument as a vector or by using choiceName + choiceValues
# if the names aren't pretty, you can pass a named vector
radioGroupButtons( # for categorical variables
  inputId = "year", # this will allow the selection to be referenced as input$year
  label = "Select Year of Vehicle", # NULL if you don't want a header
  choices = c("All", sort(unique(raw_data$year))), # can use paste, etc to make nicer display
  justified = TRUE, # will fill the width of the container it sits in (sidebar)
  status = "primary" # the color of the buttons
)

sliderInput(
  inputId = "displ", # referenced as input$displ
  label = "Select Engine Displacement:",
  value = range(raw_data$displ), # default selection, you can move the range around
  min = min(raw_data$displ),
  max = max(raw_data$displ),
  step = 0.1
)

# This will build a data frame to use throughout the dashboard. To reference this data frame, you will need to be in some reactive element ex: renderTable(), renderPlot(). This is essentially a function and so you will need to use parentheses at the end like this: renderTable(use_data())

use_data <- reactive({
  df <-
    raw_data %>%
    filter(between(displ, min(input$displ), max(input$displ)))
  
  # this if statement will filter for the selection of the radioGroupButtons
  if (input$year != "All") {
    df <- df %>% filter(year == input$year)
  }
  
  df
})
```

Column {data-width=450}
-----------------------------------------------------------------------
### Highway MPG by Engine Displacement
```{r mpg_plot}
# renderPlot is only required because we are referencing something reactive: use_data(). Otherwise we could just use ggplot(raw_data,...)
renderPlot({ 
  ggplot(use_data(), aes(displ, hwy)) +
    geom_smooth(color = "grey65") +
    geom_point(aes(color = factor(cyl))) +
    ylim(0, 45) +
    labs(color = "Cylinder") +
    theme(legend.position = "bottom")
})
```

Column
-----------------------------------------------------------------------
### Top 10 by hwy
```{r top_10, fig.height = 5.5}
#this will display a top 10 table that is filtered based on the selections above
renderTable({ # use {...} to isolate the first step (an expression)
    use_data() %>%
      arrange(desc(hwy)) %>%
      slice(1:10) %>%
      mutate(Rank = row_number()) %>%
      select(Rank, year, manufacturer, model, trans, hwy, fl, class)
  },
  # renderTable takes additional arguments after the {...}
  spacing = "xs",
  align = "l",
  bordered = T
)
```

### `r renderText({paste("Max Highway MPG by Class in Year:", input$year)})`

```{r by_class, fig.height = 4.5}
# for the header above: using `r ...`  will allow you to run R code outside of a chunk. Here the title is reactive to the year selected

renderTable({ # the curly braces allow you to make separate objects (similar to in a function)
  count_n <-
    use_data() %>%
    count(class)

  max_value <- # this could have been piped further above. It is broken into different steps here to show how it works
    use_data() %>%
    group_by(class) %>%
    summarise(hwy = max(hwy)) %>%
    ungroup()

  # and then join them together for the final product
  count_n %>%
    left_join(max_value) %>%
    arrange(desc(hwy)) %>%
    select(hwy, class, `#` = n)
  },
  spacing = "xs",
  align = "l",
  bordered = T
)
```

Your Turn
=====================================

Column 
-------------------------------------

### To-Do

The goal of this demo is to understand why each piece of code exists. Some of it is `tidyverse` manipulation, much of it is reactive `shiny` content.  Try to understand what each step does, and how to manipulate it, then work through the list below to make changes to the dashboard. 

**MPG Example**

* Convert the two table display into 2 tabs
* Change buttons to vertical display using `mpg$trans` instead of `mpg$year`
* Change slider to `mpg$hwy`

**This Page**

* use the data from `forcats::gss_cat` or `dplyr::starwars`
* Build a dashboard around demographics w/ at least 7 widgets/plots (look into `shinyWidgets`)
* Change the title of this page


### Resources

* Find more information about flexdashboards here:
    https://rmarkdown.rstudio.com/flexdashboard/using.html#overview
* The `shinyWidgets` gallery has some nice widgets for filtering:
    https://dreamrs-vic.shinyapps.io/shinyWidgets/
    The gallery isn't always available. Another overview can be found here:      
    https://dreamrs.github.io/shinyWidgets/index.html
    
Column {.tabset}
-------------------------------------
   
### Tab 1

```{r tab_1}
```   
 
### Tab 2

```{r tab_2}
```
