---
title: "Are the demographics of the two groups similar?"
author: Jake Riley
output: 
  html_document:
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
options(knitr.duplicate.label = 'allow')
```

### Prepare the data
```{r}
library(tidyverse)

df <-
  mpg %>%
  filter(year == 1999 | cty > 15) %>% 
  select(-c(displ, cty, hwy, model)) 

```

```{r, echo=FALSE}
df %>% sample_n(10)
```

### Transform to long-form

I transform the data into a long format. Each column is now stacked on top of each other in the `field` column with the original values in the `value` column.

```{r}
df_long <-
  df %>% 
  gather(field, value, -year) %>% 
  mutate(value = fct_infreq(value) %>% fct_rev())
```

```{r, echo=FALSE}
df_long %>% sample_n(10) %>% arrange(year)
```

### Visualize the differences

```{r echo = FALSE}
df_long %>% 
  ggplot(aes(value, fill = field)) +
  geom_bar() +
  coord_flip() +
  guides(fill = FALSE) +
  facet_grid(
    rows = vars(field), cols = vars(year), 
    scales = "free", space = "free"
  )
```

### Stats

Using `group_by(field) %>% do(chisq.test(year ~ value))` to perform a chi-square test by year for differences in each field (column).
```{r}
#suppressWarnings(
  chisq_res <-
    df_long %>% 
    group_by(field) %>% 
    do(broom::tidy(chisq.test(.$year, .$value))) %>% 
    ungroup() %>% 
    arrange(p.value)
#)
```
<br><br>
The biggest variation in demographics is among the **`r first(chisq_res$field)`** field and the least variation is among the **`r last(chisq_res$field)`**.
<br><br>
```{r, echo=FALSE}
chisq_res
```

```{r, echo=FALSE}
knitr::knit_exit()
```

### Appendix
```{r ref.label=knitr::all_labels(), eval=FALSE}
# https://stackoverflow.com/a/35789929/4650934
```

