---
title: "Meetup Map Demo"
author: "Jake"
date: "February 24, 2016"
output: html_document
---
**Set up workspace**
  
```{r message=FALSE, warning=F}
    library(dplyr)    # data manipulation
    library(tidyr)    # data manipulation - gather/spread functions
    library(ggplot2)  # data viz
    library(ggmap)    # data viz (maps)
    library(RCurl)    # get data from URLs
    library(knitr)    # make nicer tables in html
    
    options(scipen = 999) # turns off scientific notation
```

**Get data to map**

To assign a data table, a variable, a map, a function, etc. you use the ```<-```  symbollization. In the earlier days people used an "=" sign but this is now the accepted convention. The command ```getURL``` retrieves data from the github site. I then use piping ```%>%``` to connect ```dplyr``` commands to read the data from ```getURL``` as a csv with the ```read.csv``` command. My data has headers and use the "stringsAsFactors=FALSE" for now to save yourself from confusion. Lastly, I have the ```dplyr``` chain result in a data.frame by using ```as.data.frame()```. ```ggplot()``` uses a similar piping, but for this it is a "+" sign and adds elements to your map. 
    
```{r}
    #Get Census Data
      acsData  <- 
      getURL("https://raw.githubusercontent.com/rjake/CommonFiles/master/PHL_HousingBurden2014.csv")%>%
        read.csv(text=., header = TRUE, stringsAsFactors = FALSE)%>%  # the dot means "this data" in the pipe
        as.data.frame()                                               
    
    #Get Variables
      acsVarList  <- 
        getURL("https://raw.githubusercontent.com/rjake/CommonFiles/master/PHL_HousingBurden2014_varList.csv")%>%
        read.csv(text=., header = TRUE, stringsAsFactors = FALSE)%>%
        as.data.frame()
```

**Get census tract borders** (as .csv) & create basic maps

[This tutorial](http://www.kevjohnson.org/making-maps-in-r) shows you how to turn a shapefile to a data frame. You will need the ```rgdal``` library. In this step, the ```mutate``` command creates a new column (actually replaces the current one) that is a character version of "id" in this data.frame
    
```{r warning=F}
    #Get tract borders
      ggtract <- 
        getURL("https://raw.githubusercontent.com/rjake/FinalProject/master/ggtract.csv")%>%
        read.csv(text=., header = TRUE, stringsAsFactors = FALSE)%>%
          as.data.frame()%>%
        mutate(id=as.character(id))
```

**Create map**

You always call ```ggplot()``` and then some geom_... (ex: ```geom_polygon()```, ```geom_line()```, ```geom_point()```, ```geom_bar()```, ...).  To create a map you need to specify the x & y coordinates and what variable makes each polygon unique. You do this by using ```aes(x=, y=, group=)``` argument in the ```geom_polygon()``` command. If the colors of the polygons or colors of the borders vary based on values in the data, put them inside ```aes()```. If the color/fill is to be constant to all polygons, put them outside the ```aes()``` but still within the ```geom_polygn()``` command. Use the ```+``` sign to chain commands for creating the map together

```{r warning=F, fig.align="center"}  
      ggplot()+
          geom_polygon(data=ggtract, aes(x=long, y=lat, group=group), color="black", fill="gray95")+
          coord_fixed()+   #keeps aspect ration between x:y
          theme_nothing()  #removes gridlines, axis labels, etc
      
      #get table of census tracts with "neighborhood" IDs
        neighborhoods <- 
          getURL("https://raw.githubusercontent.com/rjake/CommonFiles/master/PhillyCentroidsWGS84.csv")%>%
          read.csv(text=., header = TRUE, stringsAsFactors = FALSE)%>%
          as.data.frame()%>%
          select(GEOID10,SECTION)%>%        #Drop all columns but these 2
          mutate(id=as.character(GEOID10))  #create column "id" as a "as.character" version of (GEOID)
      
      #Add neighborhoods to map
        Philly <- left_join(ggtract, neighborhoods, by="id") #join both data frames by the "id" column
      
      #Make map of neighborhoods
        ggplot()+
          geom_polygon(data=Philly, 
                       aes(x=long, y=lat, group=group, fill=SECTION), 
                            #fill goes inside aes() because it changes based on value in the data
                       color="black")+ 
                          #color stays outside because the black border applies to all census tracts
          coord_fixed()+
          theme_nothing(legend = T)
```


**Look at Percentage across income brackets**
Gather & Spread commands: These commands are in the ```tidyr``` library. ```gather()``` takes all of the column names, in this case, everything but 'GEOID' ```gather(..., -GEOID)``` and puts the column names in a column called Var ```gather(key=Var, ...)``` and the values in that column in an adjacent column called 'Pop' ```gather(..., value=Pop,...)```. ```spread()``` then transforms the data from long to wide where the unique values of one column, in this case 'Measure' ```spread(key=Measure,...)``` become new column headers and the values in another column, in this case Pop ```spread(..., value=Pop)``` become the values in those columns.
```{r warning=F, fig.align="center"}
    #Create a table
    acsData%>%
      gather(key=Var, value=Pop, -GEOID)%>%    #convert to long form everything but the GEOID column
      left_join(acsVarList, by="Var")%>%       #bring in VarList
      group_by(Income, Measure)%>%             #group by Income category and Measure (total or burden #s) 
      summarise(Pop=sum(Pop))%>%                  # then sum(Pop)
      spread(key=Measure, value=Pop)%>%        #Convert back to Wide form using the Measure & Pop columns
      ungroup()%>%                             #ungroup
      mutate(PctBurden=round(Burden/Total*100,2))%>%      #add a column that calculates the % of Burden/Total
      kable(align='r', format.args = list(big.mark=","))  #not necessary, used to make html/word output easier to read
        

```


**Convert to long form for ggplot2 maps facetting**

Create data to map income levels

```{r fig.width=1, warning=F}
    mapACS <- acsData%>%    #Use ACSData
      gather(key=Var, value=Pop, -GEOID)%>%     #convert to long form
      left_join(acsVarList, by="Var")%>%        #bring in variable IDs
      filter(!is.na(floorIncome))%>%            #!is.na means "is not NA"
      group_by(GEOID, Income, floorIncome, Measure)%>%  #select these columns to aggregate data then sum(Pop)
        summarise(Pop=sum(Pop))%>%                         
      ungroup()%>%                              #turn off grouping feature
      mutate(id=as.character(GEOID))%>%         #create column "id" as a "as.character" version of (GEOID)
      select(-GEOID)%>%                         #Drop the column "GEOID"
      left_join(Philly, by="id")                #Join neighborhoods labels to census tract IDs
```

***Make Maps of all income categories***

```{r warning=F, fig.align="center"}
    ggplot()+
      geom_polygon(data=Philly, aes(long,lat,group=id), color="grey80", fill="white", size=.5)+  #base map
      geom_polygon(data=mapACS, aes(x=long, y=lat, group=id, fill=Pop))+                  #map with totals
      scale_fill_gradient(low="white", high="navyblue")+           #Specifies colors for map aes(fill=Pop)
      coord_fixed()+                  #aspect ratio (x:y always fixed)
      facet_grid(Measure~Income)+     #Measure categories go down right side, Income categories across the top
      theme_nothing(legend = T)       #Removes backgrounds, axis ticks, gridlines, etc; leaves legend
```  

***Map just population  with income < $35K***

```{r warning=F}
    #Set up data
      burdenLT35 <- acsData%>%                 #these staps same as mapACS introsteps
        gather(key=Var, value=Pop, -GEOID)%>%
        left_join(acsVarList, by="Var")%>%  
        mutate(id=as.character(GEOID))%>%
        filter(!is.na(floorIncome))%>% 
        filter(floorIncome < 35)%>%            #filter where "floor income is less than 35"
        group_by(id, Measure)%>%               #group by id & measure and sum(Pop)
          summarize(Pop=sum(Pop))%>%
        ungroup()%>% 
        left_join(Philly, by="id")
```

This step puts the parts of town in a particular order. "factor" allows a categorical variable to act like an ordinal variable. The $ sign indicates the column name (```burdenLT35$SECTION``` means the "burdenLT35"" table, column "SECTION") and the command ```factor(...)``` applies to the whole column

```{r}
    burdenLT35$SECTION <- factor(burdenLT35$SECTION,
                                             levels=c("NorthWest", "FrankfordKennsington","NorthEast",
                                                      "SouthWest", "CenterCity","SouthPhilly"), ordered = TRUE)
```

Use the SECTION column to create a grid by Philly neighborhoods. Because I want the measure to double across the top I use ```facet_wrap()``` with the argument ```(... ncol=6)```

```{r warning=F, fig.align="center", fig.width=8}
    #Create map
      ggplot()+
        geom_polygon(data=burdenLT35, aes(x=long, y=lat, group=id, fill=Pop), color="grey80", size=.5)+ 
        scale_fill_gradient(low="white", high="navyblue")+
        facet_wrap(SECTION~Measure, scales="free", ncol=6)+
        theme_nothing(legend = T)+
        ggtitle("Number of People with Incomes < $35,000 who are Burdened by Housing Costs")+
        theme(plot.title = element_text(hjust = .2,size=10),    #Aligns the title
              strip.text = element_text(size = 8),              #Makes the panel titles size 8 font
              panel.background = element_rect(fill = "white"),  #makes the area behind the map white 
              aspect.ratio=1)                                   #Similar to coord_fixed when using facet_wrap
```

***TIPS***
* For both the ```dplyr``` and ```ggplot``` data, you can go line by line to see what each line does by not selecting the pipe in the last row you wish to run. For example: ```iris %>% select(Species) %>% distinct()``` You can run ```iris %>% select(Species)``` to see what it does before running the following line.

* When looking at help (ex:[StackOverflow](http://stackoverflow.com/questions/tagged/r)), people will use 'df', 'dat', 'mydata' to talk about your data. If something has brackets, like ```dat[,]``` it is a data frame and you should fill in 'dat' with your data. If it has parenthese, like ```seq()``` it is a command and you should use the command they give you. Also, those people are fast at responding and incredibly helpful. Don't be afraid to ask a newbie question. It's part of the learning process and they are happy to help.

* The example I made did not require files on my computer. If you are working with local files, be sure to use ```setwd("C:/Users/user1/Desktop")``` note that this must use *forward slashes* and not back slashes. To grab something in another folder you can move backwards with "../" in the path name, for example: ```read.csv("../example.csv")```

* Always record your process in the Source window, not the console. You can use the console to test your code but the stuff you want to repeat should be in the Source. This script is saves as a .Rmd so that I can turn it into .html, .docx, or a .pdf. You can also use .R and create all comments with the "#" sign.

* To run a line of code you can just hit ```Ctrl+Enter``` or on macs, ```Command+Enter``` as long as your cursor is in that row.

* Your commands do not have to be written all on the same line. R knows if the command is finished or not, are all parentheses closed? Is there a comma that doesn't have anything written after it. You can break up your code across lines in order to make it easier to read.

* To clear out the output of the console, hit ```Ctrl+L``` or ```Command+L```

I hope this was helpful. Please reach out if you have questions: (rjake@sas.upenn.edu)
