---
title: "Meetup Map Demo"
author: "Jake"
date: "February 24, 2016"
output: html_document
---
**Set up workspace**
  
```{r message=FALSE, warning=F}
    library(dplyr)    # data manipulation
    library(tidyr)    # data manipulation - gather/spread functions
    library(ggplot2)  # data viz
    library(ggmap)    # data viz (maps)
    library(RCurl)    # get data from URLs
    library(knitr)    # make nicer tables in html
    
    options(scipen = 999) # turns off scientific notation
```

**Get data to map**

To assign a data table, a variable, a map, a function, etc. you use the ```<-```  symbollization. In the earlier days people used an "=" sign but this is now the accepted convention. The command ```getURL``` retrieves data from the github site. I then use piping ```%>%``` to connect ```dplyr``` commands to read the data from ```getURL``` as a csv with the ```read.csv``` command. My data has headers and use the "stringsAsFactors=FALSE" for now to save yourself from confusion. Lastly, I have the ```dplyr``` chain result in a data.frame by using ```as.data.frame()```
    
```{r}
    #Get Census Data
      acsData  <- getURL("https://raw.githubusercontent.com/rjake/CommonFiles/master/PHL_HousingBurden2014.csv")%>%
        read.csv(text=., header = TRUE, stringsAsFactors = FALSE)%>%
        as.data.frame()
    
    #Get Variables
      acsVarList  <- getURL("https://raw.githubusercontent.com/rjake/CommonFiles/master/PHL_HousingBurden2014_varList.csv")%>%
        read.csv(text=., header = TRUE, stringsAsFactors = FALSE)%>%
        as.data.frame()
```

**Get census tract borders** (as .csv) & create basic maps

[This tutorial](http://www.kevjohnson.org/making-maps-in-r) shows you how to turn a shapefile to a data frame. You will need the ```rgdal``` library. In this step, the ```mutate``` command creates a new column (actually replaces the current one) that is a character version of "id" in this data.frame
    
```{r warning=F}
    #Get tract borders
      ggtract <- getURL("https://raw.githubusercontent.com/rjake/FinalProject/master/ggtract.csv")%>%
        read.csv(text=., header = TRUE, stringsAsFactors = FALSE)%>%
          as.data.frame()%>%
        mutate(id=as.character(id))
```

**Create map**

You always call ```ggplot()``` and then some geom_... (ex: ```geom_polygon()```, ```geom_line()```, ```geom_point()```, ```geom_bar()```, ...).  To create a map you need to specify the x & y coordinates and what variable makes each polygon unique. You do this by using ```aes(x=, y=, group=)``` argument in the ```geom_polygon()``` command. If the colors of the polygons or colors of the borders vary based on values in the data, put them inside ```aes()```. If the color/fill is to be constant to all polygons, put them outside the ```aes()``` but still within the ```geom_polygn()``` command. Use the ```+``` sign to chain commands for creating the map together

```{r warning=F, fig.align="center"}  
      ggplot()+
          geom_polygon(data=ggtract, aes(x=long, y=lat, group=group), color="black", fill="gray95")+
          coord_fixed()+
          theme_nothing()
      
      #get table of census tracts with "neighborhood" IDs
        neighborhoods <- getURL("https://raw.githubusercontent.com/rjake/CommonFiles/master/PhillyCentroidsWGS84.csv")%>%
          read.csv(text=., header = TRUE, stringsAsFactors = FALSE)%>%
            as.data.frame()%>%
          select(GEOID10,SECTION)%>%
          mutate(id=as.character(GEOID10))
      
      #Add neighborhoods to map
        Philly <- left_join(ggtract, neighborhoods, by="id")
      
      #Make map of neighborhoods
        ggplot()+
          geom_polygon(data=Philly, 
                       aes(x=long, y=lat, group=group, fill=SECTION), 
                            #fill goes inside aes() because it changes based on value in the data
                       color="black")+ 
                          #color stays outside because the black border applies to all census tracts
          coord_fixed()+
          theme_nothing(legend = T)
```


**Look at Percentage across income brackets**

```{r warning=F, fig.align="center"}
    #Create a table
    acsData%>%
      gather(key=Var, value=Pop, -GEOID)%>%    #convert to long form
      left_join(acsVarList, by="Var")%>%       #bring in VarList
      group_by(Income, Measure)%>%             #group by Income/Measure and sum(Pop)
      summarise(Pop=sum(Pop))%>%
      spread(key=Measure, value=Pop)%>%        #Convert back to Wide form
      ungroup()%>%                             #ungroup
      mutate(PctBurden=round(Burden/Total*100,2))%>%      #add a column that calculates the % of Burden/Total
      kable(align='r', format.args = list(big.mark=","))  #not necessary, used to make html/word output easier to read
        

```


**Convert to long form for ggplot2 maps facetting**

Create data to map income levels

```{r fig.width=1, warning=F}
    mapACS <- acsData%>%    #Use ACSData
      gather(key=Var, value=Pop, -GEOID)%>%     #convert to long form
      left_join(acsVarList, by="Var")%>%        #bring in variable IDs
      filter(!is.na(floorIncome))%>%            #!is.na means "is not NA"
      group_by(GEOID, Income, floorIncome, Measure)%>%  #select these columns to aggregate data then sum(Pop)
        summarise(Pop=sum(Pop))%>%                         
      ungroup()%>%                              #turn off grouping feature
      mutate(id=as.character(GEOID))%>%         #create column called "id", that is a "as.character" version of the FIPS code (GEOID)
      select(-GEOID)%>%                         #Drop the column "GEOID"
      left_join(Philly, by="id")                #Join neighborhoods labels to census tract IDs
```

***Make Maps of all income categories***

```{r warning=F, fig.align="center"}
    ggplot()+
      geom_polygon(data=Philly, aes(long,lat,group=id), color="grey80", fill="white", size=.5)+  #base map
      geom_polygon(data=mapACS, aes(x=long, y=lat, group=id, fill=Pop))+                  #map with totals
      scale_fill_gradient(low="white", high="navyblue")+           #Specifies colors for map aes(fill=Pop)
      coord_fixed()+                  #aspect ratio (x:y always fixed)
      facet_grid(Measure~Income)+     #Measure categories go down right side, Income categories across the top
      theme_nothing(legend = T)       #Removes backgrounds, axis ticks, gridlines, etc; leaves legend
```  

***Map just population  with income < $35K***

```{r warning=F}
    #Set up data
      burdenLT35 <- acsData%>%                 #these staps same as mapACS introsteps
        gather(key=Var, value=Pop, -GEOID)%>%
        left_join(acsVarList, by="Var")%>%  
        mutate(id=as.character(GEOID))%>%
        filter(!is.na(floorIncome))%>% 
        filter(floorIncome < 35)%>%            #filter where "floor income is less than 35"
        group_by(id, Measure)%>%               #group by id & measure and sum(Pop)
          summarize(Pop=sum(Pop))%>%
        ungroup()%>% 
        left_join(Philly, by="id")
```

This step puts the parts of town in a particular order. "factor" allows a categorical variable to act like an ordinal variable. The $ sign indicates the column name (```burdenLT35$SECTION``` means the "burdenLT35"" table, column "SECTION") and the command ```factor(...)``` applies to the whole column

```{r}
    burdenLT35$SECTION <- factor(burdenLT35$SECTION,
                                             levels=c("NorthWest", "FrankfordKennsington","NorthEast",
                                                      "SouthWest", "CenterCity","SouthPhilly"), ordered = TRUE)
```

Use the SECTION column to create a grid by Philly neighborhoods. Because I want the measure to double across the top I use ```facet_wrap()``` with the argument ```(... ncol=6)```

```{r warning=F, fig.align="center"}
    #Create map
      ggplot()+
        geom_polygon(data=burdenLT35, aes(x=long, y=lat, group=id, fill=Pop), color="grey80", size=.5)+ 
        scale_fill_gradient(low="white", high="navyblue")+
        coord_fixed()+
        facet_wrap(Measure~SECTION, scales="free", ncol=6)+
        theme_nothing(legend = T)+
        ggtitle("Number of People with Incomes < $35,000 who are Burdened by Housing Costs")+
        theme(plot.title = element_text(hjust = .2,size=10),
              panel.background = element_rect(fill = "white"))
```
