---
title: "How Many Observations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, comment = "")
```

```{r}
library(tidyverse)
options(scipen = 999)
```

```{r}
goal <- 0.60

get_zscore <-
    function(x){
        qnorm(1-(1-x)/2) %>% round(2)
    }

get_zscore(.995)

find_n <-
    function(p, z, se){
        #p <- 0.1; n <- 200; z <- get_zscore(0.8); se <- z*sqrt( (p*(1-p)) / n) #0.041771
        ( sqrt(p*(1-p)) / (se/z) )^2  %>% floor()
    }

find_n_pdiff <-
    function(p, z, se){
        #p <- 0.1; n <- 200; z <- get_zscore(0.8); 
        #se <- z*sqrt( (var_goal) + (samp_var)/n) 
        var_goal <- sqrt( (.6*(1-.6)) / 200 )
        var_sample <- (p*(1-p))
    
        ( (sqrt(p*(1-p)) / (se/z)) - var_goal)^2  %>% floor()
    }

find_n_pdiff(.55, 1.96, 0.05)
```

```{r}
get_values <-
    tibble(p = seq(0.02, goal - 0.02, by = 0.02)) %>%
    expand(p, 
           conf = c(90, 95, 99, 99.5)) %>%
    mutate(z = get_zscore(conf/100),
           dist = round(goal-p, 2)) %>%
    mutate(min_n = find_n(p, z, dist),
           min_n_pdiff = find_n_pdiff(p, z, dist)) %>%
    filter(min_n > 1) %>% 
    mutate_at(vars(-c(min_n, min_n_pdiff)), as.factor)

plot_colors <- c(paste0("grey", (3:1/4)*100), "blue")

p_hat <- expression(hat(p))

plot_table <-
    ggplot(get_values, aes(dist, conf)) +
    geom_tile(fill = "white") +
    geom_text(aes(label = min_n, color = z)) +
    #geom_text(aes(y = "dist.", label = dist)) +
    geom_tile(aes(y = "p-hat"), fill = "grey95") +
    geom_text(aes(y = "p-hat", label = p)) +
    scale_color_manual(values = plot_colors) +
    theme(#panel.background = element_rect(fill = "white"),
          legend.position = "none") +
    labs(x = paste0("Distance from Goal (", goal, ")"),
         y = "conf.")

#plot_table

plot_labs <-
    get_values %>% 
    filter(as.numeric(z) == max(as.numeric(z)))

plot_distribution <-
    ggplot(get_values, aes(dist, min_n)) +
    geom_line(aes(group = p), size = 3, color = "white") +
    geom_point(aes(color = z), size = 3) +
    geom_text(data = plot_labs, aes(y = min_n * 1.3, label = min_n), hjust = 0.5) +
    scale_color_manual(values = plot_colors) +
    scale_y_continuous(trans = "log", breaks = c(1, 10, 50, 100, 500, 1000, 5000)) +
    theme(panel.grid.minor = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.x = element_blank(), 
          #plot.subtitle = element_text(hjust = 1),
          legend.position = "none") +
    labs(y = "Minimum Obs. Needed (at log intervals)")

#plot_distribution
```

### Minimum N Needed Given Distance of Sample Mean from Goal (0.6)

With a goal of 0.6, a sample proportion of 0.5 would be 0.1 from the goal. Holding the SE constant at 0.1 and the conf. at 99.5%, the minimum observations needed to reject the null hypothesis that the means are equal is 197.

Formula: \(N =  \left(\frac{\sqrt{var}} {(SE/z)}\right)^2\)
&nbsp;&nbsp;&nbsp;&nbsp;
Example: \(197 = \left(\frac{\sqrt{0.5(1-0.5)}} {(0.1/2.81)}\right)^2\)
&nbsp;&nbsp;&nbsp;&nbsp;
Derived using: \(SE\ 0.1 = 2.81 * \sqrt{\frac{0.5(1-0.5)} {197}}\)

```{r echo = F, fig.height = 8, fig.width= 12}
gridExtra::grid.arrange(plot_distribution, plot_table, 
                        ncol = 1, heights = c(3,1))
```

## Questions
* should I be using this instead [Difference between 2 proportions](https://www.dummies.com/education/math/statistics/how-to-estimate-the-difference-between-two-proportions/) - looks like it returns the same results 
* should the min n be adjusted to allow for a sample size that results in that distance? For example, 0.52 from 0.6 is 0.08. 0.08/0.6 is 13%. In order to get 13%, there would need to be at least 2/25 charts
    * `MASS::fractions(1:29/50)`
    * `MASS:::.rat(1:29/50)$rat`
    
    
    
```{r echo = F, eval = F}
#title = "Minimum N Needed Given Distance of Sample Mean from Goal (0.6)",
    #     subtitle = expression(atop("With a goal of 0.6, a sample proportion of 0.5 would be 0.1 from the goal. Holding the SE constant at 0.1 and the conf. at 99.5%, the min obs needed to reject the null hypothesis that the means are equal is 197.",
    #                                " N = " ( sqrt(var) / (SE/z) )^2~
    #                                ", example: 197 = " ( sqrt(0.5*(1-0.5)) / (0.1/2.81) )^2 ~ ", derived using SE of 0.1 = " ~ 2.81 ~"*"~sqrt( 0.5*(1-0.5) / 197))),
```

